---
title: "mixWAS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mixWAS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
### Load mixWAS library
library(mixWAS)
```

To demonstrate how to use `mixWAS`, we'll first load in some sample data that was used in simulations. In particular, the following is a single dataset from Simulation 1 in the paper.

* 8 phenotypes (1-4 continuous, 5-8 binary)
* 8 phenotypes at site 1 and 5, 7 at site 2, 6 at site 3, and 5 at site 4
* 5 sites
* 1000 subjects per site
* 6 covariates 
* PC1, PC2, PC3, PC4 represent 4 principle components from a PCA analysis
* age_centered represents subject age (relative to some mean)
* gender = binary gender covariate

In this simulated dataset $\beta_1 = \beta_5 = 0.01872$ and $\beta_2 = \beta_3 = \beta_4 = \beta_6 = \beta_7 = \beta_8 = 0$.

```{r}
### Load example data from mixWAS package
data(example_data)
```


## mixWAS when all data is available to central analyst

In this example, we have access to all sites data at once. If this is the case, `mixWAS` can be called directly as follows
```{r}
mixWAS(snps = example_data$snps,
       covariates = example_data$covariates,
       phenotypes = example_data$phenotypes)
```

* `p_score`: $p$-value for score test against dense alternatives
* `p_acat`: ACAT $p$-value against sparse alternatives
* `p_snp`: Overall SNP $p$-value for detecing multi-phenotype association

If we had wanted to run this in parallel across sites, we could do so as follows.

```{r, warning = F, message = F}
library(furrr)
plan(future::multicore(workers = parallel::detectCores() - 1))
options(future.globals.maxSize = 8 * 1024^3)

mixWAS(snps = example_data$snps,
       covariates = example_data$covariates,
       phenotypes = example_data$phenotypes,
       parallel_sites = T)
```

If we had wanted intermediate outputs rather than $p$-values, we could run `mixWAS` as follows
```{r}
### mixWAS w/ intermediate components
components <- 
  mixWAS(snps = example_data$snps,
         covariates = example_data$covariates,
         phenotypes = example_data$phenotypes,
         return_p = F)

names(components)
```

* `score`: combined score vector
* `V`: Variance matrix for score vector
* `V_inv`: Inverse of variance matrix for score vector
* `z`: standardized score vector (e.g. $V^{-1/2} S$ where $S$ is the score vector).
* `q`: # of phenotypes

`mixWAS` $p$-values can be recovered from these intermediate components using `run_hypothesis_tests`.

```{r}
### Recover p-values from intermediate output
run_hypothesis_tests(score = components$score,
                     V_inv = components$V_inv,
                     z = components$z,
                     q = components$q)

```

Note that we did not need to specify the inputs `phenotype_index`, `covariate_map`, or `types` as they can be inferred. For user reference, here is the result of `mixWAS` figuring out each of these arguments.

```{r}
### Phenotype Index
### (1) Get names of all phenotypes
all_names <- unique(unlist(purrr::map(example_data$phenotypes, ~colnames(.x))))

### (2) Get numeric index of which phenotypes are at each site
phenotype_index <- 
  purrr::map(example_data$phenotypes, 
             ~which(all_names %in% colnames(.x)))
phenotype_index

### Covariate Data Types
### (1) is_binary helper function
is_binary <- function(x) {
  x <- x[!is.na(x)]
  return(all(x == 0 | x == 1))
}

### (2) Determine data types (never infers 'count')
type_list <-
  purrr::map(example_data$phenotypes, 
             ~ifelse(apply(.x, 2, is_binary), 'binary', 'continuous'))
type_list

### Covariate Map for this example
df_map <- dplyr::tibble('variable' = c('PC1', 'PC2', 'PC3', 'PC4', 
                                       'age_centered', 'gender'),
                        'phenotype' = 'all')


### Running w/ extra arguments to show same results
mixWAS(snps = example_data$snps,
       covariates = example_data$covariates,
       phenotypes = example_data$phenotypes,
       phenotype_index = phenotype_index,
       types = type_list,
       covariate_map =  purrr::map(1:8, ~df_map))

```


## mixWAS when each site needs to be processed seperately

In practice however, it's possible that individual data can not be combined across sites or even stored at the same time. When this is the case, we can run `mixWAS_single_site` on each site, and combined results, as follows. Below we show running `mixWAS_single_site` on each site a single machine but of course it could run seperately with these outputs being shared and later combined via `combine_site_results`.


```{r}
### Run mixWAS_single_site 1 at a time
num_sites <- 5
site_components <-
  map(1:num_sites, ~mixWAS_single_site(snps = example_data$snps[[.x]],
                                       phenotypes = example_data$phenotypes[[.x]],
                                       covariates = example_data$covariates[[.x]]))

### Combine results at "central analyst"
combine_site_results(mixWAS_components = site_components,
                     q = 8,
                     phenotypes = example_data$phenotypes)

```
